\begin{center}
	VITA\par
\end{center}

\subsection*{Education}
%
Laurea Magistralis in Economics, \textit{cum laude} \hfill  2010-2012 \\
 University of Bologna, Bologna, Italy\\
Laurea Triennale in Economia e Commercio, \textit{cum laude} \hfill 2007-2010\\
 University of Padova, Padua, Italy\\

\subsection*{Research Interests}
Primary: Development Economics and Behavioral Economics\\
Secondary: Political Economy and Household Finance\\

\subsection*{Working Papers}
``Moral Incentives in Credit Card Debt Repayment: Evidence from a Field Experiment", with Leonardo Bursztyn, Daniel Gottlieb, and Martin Kanz.\\
``Status Goods: Experimental Evidence from Platinum Credit Cards'', with Leonardo Bursztyn, Bruno Ferman, Martin Kanz, and Guatam Rao.\\
``From Extreme to Mainstream: How Social Norms Unravel'', with Leonardo Bursztyn and Georgy Egorov.\\
``Corruption and Legislature Size: Evidence from Bazil", with Diogo Britto.\\

\subsection*{Research Grants}
Center for Global Management (CGM) UCLA Anderson PhD Global Research Award 2016-2017 (\$4,000). Project: ``Preference Based Home Bias: the Effect of Cultural Proximity on Loan Repayment (Philippines).''\\
Center for Global Management (CGM) UCLA Anderson PhD Global Research Award 2015-2016 (\$4,000). Project: ``Speaking the Same Language: the Effect of Cultural Proximity on Loan Repayment (Kenya)."\\
Abdul Latif Jameel Poverty Action Lab (J-PAL) Governance Initiative Pilot Study Research Grant (\$49,050). Project: ``Participatory Budgeting and Urban Services: Mohalla Sabhas in Delhi."\\
Center for Global Management (CGM) UCLA Anderson PhD Global Research Award 2014-2015 (\$4,000). Project: ``Speaking the Same Language: the Effect of Cultural Proximity on Loan Repayment (Indonesia)."\\

\subsection*{Fellowships and Awards}
%
Best Paper Award, KFUPM Islamic Banking and Finance Research Conference Riyadh \hfill 2016\\
Dissertation Year Fellowship, UCLA Graduate Division \hfill 2016-2017\\
Dean's Award for early Advancement to Candidacy, UCLA Anderson \hfill 2015-2016\\
Anderson Fellowship, UCLA Anderson \hfill 2012-2016\\
Laurea Magistrale in Economics Best Student Award, University of Bologna \hfill 2012\\

\subsection*{Seminars and Invited Presentations}
%P
Pacific Conference for Development Economics (Pac Dev 2017). UC Riverside. \hfill 2017 \\
Bocconi Baffi Carefin Annual International Banking Conference. Bocconi University. \hfill 2016 \\
Microeconometric and Public Policy Seminar. The University of Sydney. \hfill 2016 \\
Warwick Economics PhD Conference 2016. The University of Warwick. \hfill 2016\\

\subsection*{Teaching Experience}
%
TA for Managerial Economics (MBA students, Nico Voigtlaender). UCLA. \hfill 2013-2017\\
TA for Statistics (MA students, Paola Bortot). University of Bologna. \hfill 2011-2012\\
TA for Mathematics (BA students, Enrico Bernardi). University of Bologna. \hfill 2010-2011\\

\subsection*{Employment}
Short Term Consultant, The World Bank. \hfill 2017\\
RA for Leonardo Bursztyn, UCLA Anderson. \hfill 2013-2017\\
Short Term Consultant, The World Bank. \hfill 2014-2015\\
RA for Rocco Macchiavello and Chris Woodruff, University of Warwick and IPA-Bangladesh. \hfill 2012\\
RA for Andrea Ichino, Margherita Fort and Giulio Zanella, University of Bologna. \hfill 2011-2012\\
RA for Erich Battistin, Sascha Becker, and Luca Nunziata, University of Padova. \hfill 2010\\

\clearpage
